Timid tides or tall tails? \\ Investigating the outer reaches of the Sculptor and Ursa Minor dwarf galaxies

### 2025-04-15 thesis/data.tex
thesis updates

Gaia provides unprecedented accuracy in proper motions and magnitudes.
As such, Gaia data is uniquely excellent to produce low-contamination
samples of likely member stars belonging to satellites. Here, we breifly
describe J+24's membership estimation and discuss how this informs our
observational knoledge of each galaxies density profile. In general,
J+24 use a Bayesian framework incorporating proper motion (PM,
colour-magnitude diagram (CMD), and spatial information to determine the
probability that a given star belongs to the satellite. J+24 extends
\citet{MV2020a} (see also \citet{pace+li2019}, etc.).

J+24 select stars initially from Gaia satisfying:

\begin{itemize}
\tightlist
\item
  High quality astrometry (\texttt{ruwe\ \textless{}=\ 1.3})
\item
  3\(\sigma\) consistency of measured parallax with dwarf distance +
  uncertainty (near zero with \citet{lindegren+2018} zero-point
  correction)
\item
  Absolute RA and Dec proper motions less than
  10\(\,{\rm mas\ yr^{-1}}\)
\item
  Good photometry
\item
  No colour excess (\citet{lindegren+2018} equation C.2)
\item
  G \textgreater{} 22 and less than 5\(\sigma\) above TRGB, and between
  -0.5 and 2.5 in Bp - Rp.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  CMD: The CMD is from Padova \citep{girardi+2002} isochrone at 12 Gyr
  with a colour width of 0.1 mag plus the Gaia colour uncertanty and the
  observed metallicity of the dwarf. The HB is modelled as a constant
  magnitude extending blue of the CMD with a 0.1 mag width plus the mean
  colour error. A likelihood map is constructed by sampling the distance
  modulus in addition to the CMD width, taking the maximum of RGB and HB
  likelihoods.
\item
  Spatial: A single exponential
  (\(\Sigma \propto e^{R_{\rm ell} / R_s}\)) using mean values of the
  structural parameters (\textbf{verify}). For Scl and UMi, this is
  instead a double exponential
  \(\Sigma_\star \propto e^{-R/R_s} + B\,e^{-R/R_{\rm outer}}\) where
  the inner exponential remains fixed and structural parameters are not
  accounted for. Structural parameter uncertainties are not accounted
  for.
\item
  PM. A bivariate gaussian with variance and covariance equal to each
  star's proper motions. Each star's proper motions uncertainty are
  assumed to be the dominant uncertainty.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  CMD : Constructed as a KDE using the other quality-selected stars
  outside of \(5R_h\) in the catalogue. The width of each star is
  projected as a bivariate Gaussian.
\item
  PM: same as CMD except in PM space.
\item
  Spatial: a constant likelihood.
\end{itemize}

Note that each likelihood map is normalized over the respective
parameter space. In order to represent the difference in frequency of
background and forground stars, \(f_{\rm sat}\) represents the field
fraction of member stars.

@@

\begin{itemize}
\tightlist
\item
  Systemic proper motions \(\mu_\alpha\), \(\mu_\delta\). Prior: within
  5\(\sigma\) of single component case w/ systematic uncertainties.
  Single component prior is based on MV2020 (RV members / expected
  velocity dispersion of halo at distance?)
\item
  \(f_{\rm sat}\) density normalization (uniform between 0 and 1)
\item
  Spatial component parameters \(B\) and \(R_{\rm outer}\) for extended
  profiles (Scl and UMi here.)
\end{itemize}

The median parameters from the simulation are then used to calculate the
final \(P_{\rm sat}\) values we use here.

@@

\begin{figure}
\centering
\includegraphics{figures/scl_selection.pdf}
\caption[Sculptor selection criteria]{The selection criteria for Scl
members. Probable members (2-component) are orange, and all field stars
(satisfying quality criteria) are in light grey. \textbf{Top:} Tangent
plane. \textbf{Bottom left:} Colour magnitude diagram. \textbf{Bottom
right:} Proper motion.}\label{fig:sculptor_selection}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/umi_selection.pdf}
\caption[Ursa Minor Selection]{Similar to
fig.~\ref{fig:sculptor_selection} except for Ursa Minor. UMi features a
very extended density profile with some stars \textasciitilde{} 6\(R_h\)
including a RV member. UMi is also highly elliptical compared to other
classical dwarfs.}\label{fig:umi_selection}
\end{figure}

Figure: Similar to fig.~\ref{fig:sculptor_selection} except for Fornax.
Note that we only use one-component probabilities here. Fornax is the
quintessential unperturbed dwarf galaxy.

\subsubsection{Searches for tidal tails}\label{searches-for-tidal-tails}

\begin{itemize}
\item
  There are no apparent overdensities in the PM \& CMD only selected
  stars to suggest the presence of a tidal tail
\item
  We have tried selective matched filter
\item
  This means that at least at the level of where the background density
  dominates, we can exclude models which produce tidal tails brighter
  than a density of
  \(\Sigma_\star \approx 10^{-2}\,\text{Gaia-stars\ arcmin}^{-2} \approx 10^{-6} \, {\rm M_\odot\ kpc^{-2}}\)
  (TODO assuming a distance of \ldots{} and stellar mass of \ldots).
\end{itemize}

@@

To derive density profiles, we use 0.05 dex bins in log radius (i.e.~the
bins are derived from 10\^{}(minimum(logR):0.05:maximum(logR))). The
density in each bin is then (from Poisson statistics) \[
\Sigma_b = N_{\rm memb} / A_{\rm bin} \pm \sqrt{N_{\rm memb}} / A_{\rm bin}
\] where \(N_{\rm memb}\) is the number of members in the bin and
\(A_{\rm bin}\) is the area of the bin. As discussed below, these
uncertainties underrepresent the true uncertainty on multiple accounts.
We retain Poisson errors for simplicity here.

@@

\begin{figure}
\centering
\includegraphics{figures/scl_density_methods.pdf}
\caption[Sculptor density profiles]{The density profile of Sculptor for
different selection criteria. \emph{probable members} selects stars with
PSAT \textgreater{} 0.2 considering PM, CMD, and spatial, \emph{CMD+PM}
select stars more likely to be members according to CMD and PM only,
\emph{all} selects any high quality star, and \emph{BG subtracted} is
the background-subtracted density derived from high-quality
stars.}\label{fig:scl_observed_profiles}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/umi_density_methods.pdf}
\caption[Ursa Minor density profiles]{Similar to
fig.~\ref{fig:scl_observed_profiles} except for Ursa
Minor.}\label{fig:umi_observed_profiles}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/fornax_density_methods.pdf}
\caption[Fornax density profiles]{Similar to
fig.~\ref{fig:scl_observed_profiles} except for
Fornax.}\label{fig:fornax_observed_profiles}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{figures/scl_umi_fornax_exp_fit.pdf}
\caption[Classical dwarf density profiles]{The density profiles of
Sculptor, Ursa Minor, and Fornax compared to Exp2D and Plummer density
profiles. Dwarf galaxies are scaled to the same half-light radius and
density at half-light radius (fit from the inner 3 scale radii
exponential recursively. )}\label{fig:classical_dwarfs_densities}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{figures/scl_density_methods_extra.pdf}
\caption[Density profiles]{Density profiles for various assumptions for
Sculptor. PSAT is our fiducial 2-component J+24 sample, circ is a
2-component bayesian model assuming circular radii, simple is the series
of simple cuts described in Appendix ?, bright is the sample of the
brightest half of stars (scaled by 2), DELVE is a sample of RGB stars
(background subtracted and rescaled to
match).}\label{fig:sculptor_observed_profiles}
\end{figure}

@@

>>>>

Gaia provides unprecedented accuracy in proper motions and magnitudes.
As such, Gaia data is uniquely excellent to produce low-contamination
samples of likely member stars belonging to satellites. Here, we breifly
describe J+24's membership estimation and discuss how this informs our
observational knoledge of each galaxies density profile. In general,
J+24 use a Bayesian framework incorporating proper motion (PM),
colour-magnitude diagram (CMD), and spatial information to determine the
probability that a given star belongs to the satellite. J+24 extends
\citet{MV2020a} (see also \citet{pace+li2019}, \citet{battaglia+2022},
\citet{pace+erkal+li2022}, etc.).

J+24 select stars initially from Gaia within a 2--4 degree circular
region centred on the dwarf satisfying:

\begin{itemize}
\tightlist
\item
  Solved parallax, proper motions, colour, and magnitudes.
\item
  High quality astrometry (\texttt{ruwe\ \textless{}=\ 1.3})
\item
  3\(\sigma\) consistency of measured parallax with dwarf distance +
  uncertainty (typically near zero; with \citet{lindegren+2018}
  zero-point correction).
\item
  Absolute RA and Dec proper motions less than
  10\(\,{\rm mas\ yr^{-1}}\) (corresponding to tangental velocities of
  \(\gtrsim 500\) km/s at distances larger than 10 kpc.)
\item
  No colour excess (\citet{lindegren+2018} equation C.2)
\item
  G \textless{} 22 and less than 5\(\sigma\) above TRGB, and between
  -0.5 and 2.5 in Bp - Rp.
\end{itemize}

Photometry is dereddened using \citet{schlegel+1988} extenction.

@@

\begin{itemize}
\tightlist
\item
  CMD: The CMD is the lowest metallicity isochrone from Padova
  \citep{girardi+2002} with age 12 Gyr with a colour width of 0.1 mag
  plus the Gaia colour uncertainty at each magnitude. The HB is modelled
  as a constant magnitude extending blue of the CMD. The HB magnitude is
  the mean magnitude of HB stars from most metal poor isochrone with a
  0.1 mag width plus the mean colour error. A likelihood map is
  constructed by sampling the distance modulus in addition to the CMD
  width, taking the maximum of RGB and HB likelihoods.
\item
  Spatial: A single exponential
  (\(\Sigma \propto e^{R_{\rm ell} / R_s}\)) accounting for structural
  uncertainty (sampled over position angle, ellipticity, and half light
  radius).
\item
  Alternative spatial: For Scl and UMi, this is instead a double
  exponential
  \(\Sigma_\star \propto e^{-R/R_s} + B\,e^{-R/R_{\rm outer}}\) where
  the inner exponential remains fixed. Structural parameter
  uncertainties are not accounted for.
\item
  PM. A bivariate gaussian with variance and covariance equal to each
  star's proper motions. Each star's proper motions uncertainty are
  assumed to be the dominant uncertainty.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  CMD : Constructed as a density map using the other quality-selected
  stars outside of \(5R_h\) in the catalogue. The map is a sum of
  bivariate gaussians for each star with standard deviations based on
  the Gaia uncertainties.
\item
  PM: same as CMD except in PM space.
\item
  Spatial: a constant likelihood.
\end{itemize}

Note that each likelihood map is normalized over the respective 2D
parameter space. In order to represent the difference in frequency of
background and forground stars, \(f_{\rm sat}\) represents the field
fraction of member stars.

@@

\begin{itemize}
\tightlist
\item
  Systemic proper motions \(\mu_\alpha\), \(\mu_\delta\). Single
  component prior is same as \citet{MV2020}: a normal distribution with
  mean 0 and standard deviation 100 km/s. If 2-component spatial,
  instead is a uniform distribution spanning 5\(\sigma\) of single
  component case w/ systematic uncertainties.
\item
  \(f_{\rm sat}\) density normalization. Prior is a uniform distribution
  between 0 and 1.
\item
  Spatial component parameters \(B\) is uniform from 0-1 and
  \(R_{\rm outer}\) is uniform and greater than \(R_s\) for extended
  profiles (Scl and UMi here.)
\end{itemize}

The mode of each parameter from the MCMC are then used to calculate the
final \(P_{\rm sat}\) values we use here.

@@

In figures fig.~\ref{fig:sculptor_selection},
fig.~\ref{fig:umi_selction}, we illustrate the resulting samples from
the algorithm. In each case, each criteria plays a roll: proper motions
are centred around the dwarf systemic motion, the CMD is well defined,
and stars only within a few \(R_h\) are included. We also plot the RV
members found in general and in \citet{sestito+2024},
\citet{sestito+2024b}.

The tangent plane, \(\xi\), \(\eta\), is the projection that

We also illustrate the approximate result of removing the spatial
component from the likelihood. We define the CMD+PM selection as stars
satisfying \[
{\cal L}_{\rm CMD,\ sat}\ {\cal L}_{\rm PM,\ sat} > {\cal L}_{\rm CMD,\ bg}\ {\cal L}_{\rm PM,\ bg}
\]

These stars appear similar to the fiducial (probable members) sample,
but instead also appear as an approximately uniform distribution across
the entire field. This illustrates the approximate background of stars
which may be confused as members. Additionally, since there is no clear
spatial structure in the CMD+PM sample, it is unlikely that there are
tidal tails detectable with Gaia. Not shown here, we also try a variety
of simpler, absolute cuts and thresholds, finding no extended structure
beyond what is detected in J+24.

This means that at least at the level of where the background density
dominates, we can exclude models which produce tidal tails brighter than
a density of
\(\Sigma_\star \approx 10^{-2}\,\text{Gaia-stars\ arcmin}^{-2} \approx 10^{-6} \, {\rm M_\odot\ kpc^{-2}}\)
(TODO assuming a distance of \ldots{} and stellar mass of \ldots).

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_selection.pdf}}
\caption[Sculptor selection criteria]{The selection criteria for Scl
members. Probable members (2-component) are orange, and all field stars
(satisfying quality criteria) are in light grey. \textbf{Top:} Tangent
plane. \textbf{Bottom left:} Colour magnitude diagram. \textbf{Bottom
right:} Proper motion.}\label{fig:sculptor_selection}
\end{figure}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/umi_selection.pdf}}
\caption[Ursa Minor Selection]{Similar to
fig.~\ref{fig:sculptor_selection} except for Ursa Minor. UMi features a
very extended density profile with some stars \textasciitilde{} 6\(R_h\)
including a RV member. UMi is also highly elliptical compared to other
classical dwarfs.}\label{fig:umi_selection}
\end{figure}

@@

To derive density profiles, we use 0.05 dex bins in log radius (i.e.~the
bins are derived from 10\^{}(minimum(logR):0.05:maximum(logR))). The
density in each bin is then (from Poisson statistics) \[
\Sigma_b = N_{\rm memb} / A_{\rm bin} \pm \sqrt{N_{\rm memb}} / A_{\rm bin}
\] where \(N_{\rm memb}\) is the number of members in the bin and
\(A_{\rm bin}\) is the area of the bin's annulus in 2D. As discussed
below, these uncertainties underrepresent the true uncertainty on
multiple accounts. We retain Poisson errors for simplicity here.

@@

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_density_methods.pdf}}
\caption[Sculptor density profiles]{The density profile of Sculptor for
different selection criteria. \emph{probable members} selects stars with
PSAT \textgreater{} 0.2 considering PM, CMD, and spatial, \emph{CMD+PM}
select stars more likely to be members according to CMD and PM only,
\emph{all} selects any high quality star, and \emph{BG subtracted} is
the background-subtracted density derived from high-quality
stars.}\label{fig:scl_observed_profiles}
\end{figure}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/umi_density_methods.pdf}}
\caption[Ursa Minor density profiles]{Similar to
fig.~\ref{fig:scl_observed_profiles} except for Ursa
Minor.}\label{fig:umi_observed_profiles}
\end{figure}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/fornax_density_methods.pdf}}
\caption[Fornax density profiles]{Similar to
fig.~\ref{fig:scl_observed_profiles} except for
Fornax.}\label{fig:fornax_observed_profiles}
\end{figure}

@@

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_umi_fornax_exp_fit.pdf}}
\caption[Classical dwarf density profiles]{The density profiles of
Sculptor, Ursa Minor, and Fornax compared to Exp2D and Plummer density
profiles. Dwarf galaxies are scaled to the same half-light radius and
density at half-light radius (fit from the inner 3 scale radii
exponential recursively. )}\label{fig:classical_dwarfs_densities}
\end{figure}

@@

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_density_methods_extra.pdf}}
\caption[Density profiles]{Density profiles for various assumptions for
Sculptor. PSAT is our fiducial 2-component J+24 sample, circ is a
2-component bayesian model assuming circular radii, simple is the series
of simple cuts described in Appendix ?, bright is the sample of the
brightest half of stars (scaled by 2), DELVE is a sample of RGB stars
(background subtracted and rescaled to
match).}\label{fig:sculptor_observed_profiles}
\end{figure}

@@

\section{Radial velocity modeling}\label{radial-velocity-modeling}

For both Sculptor and Ursa Minor, we construct samples of radial
velocity measurements and cross match each sample to produce estimates
of the total velocity dispersion and line of sight velocity for each
galaxy.

\subsection{Methodology}\label{methodology}

Our RV methodology is

\begin{itemize}
\tightlist
\item
  Cross match all catalogues to J+24 Gaia members. If Gaia ID's are
  included, we use these. Otherwise, we match to the nearest star within
  2 arcseconds.
\item
  We combine the mean RV measurement from each study using the
  inverse-variance weighted mean and standard uncertainty.
\end{itemize}

\[
v_i = \frac{1}{\sum w}\sum_j w\ v_{i, j} \\
\delta v_i = \sqrt{\frac{1}{\sum w^2}}
\]

where \(w_i = 1/s_i^2\), and we estimate the inter-study standard
deviation \[
s_{v}^2 = \frac{1}{\sum w} \sum w (v_{i, j} - v_i)^2
\] Note that we do not include the standard deviation if a source does
measure the velocity multiple times.

If the total standard deviation is greater than five times the velocity
standard uncertainty (reduced), then we remove the star since either the
star is a binary or contains erroneous measurements or other
uncertainties. Note that since stars typically only have very few
measurements, these choices are highly approximate and likely limited.

The combined RV likelihood is then \[
{\cal L} = {\cal L}_{\rm space} {\cal L}_{\rm CMD} {\cal L}_{\rm PM} {\cal L}_{\rm RV}
\] where \[
{\cal L}_{\rm RV} = N(\mu_{v}, \sigma_{v}^2 + (\delta v_i)^2)
\] where \(\mu_v\) and \(\sigma_v\) are the systemic velocity and
dispersion, and \(\delta v_i\) is the individual measurement
uncertainty. Typically, the velocity dispersion will dominate the
uncertainty budget here.

Similar to above, we retain stars with the resulting membership
probability of greater than 0.2.

Finally, we need to correct the coordinate frames. We need a proper
motion estimate of each star. Since most stars have Gaia measured proper
motions, but the uncertainties on each proper motion measurement are
large, we combine the system and individual star's proper motion by
weighting the systemic proper motion by the z-score of the consistency
of the measurement with the systemic proper motion. \[
\mu_{*}' = \frac{1}{w_\star + w_{\rm sys}}\left(\mu_{\star}w_\star + \mu_{\rm sys} w_{\rm sys}\right)
\] where we model \[
{w_{\rm sys}} = P_{\rm sat}(\delta \mu ^2 + \sigma_{v, tan}^2)^{-1}
\] and the weight of the star is just the inverse variance weight plus
the gaia systematics.

The corrected radial velocities are then \[
v_z = v_{\rm los, gsr}\cos\theta  + v_{\rm tan, R} \sin\theta
\] where
\(v_{\rm tan, R} = d(\mu_{\alpha*}\cos\theta + \mu_\delta \sin \theta)\)
is the radial component of the proper motion with respect to the centre
of the galaxy. Since \(\theta \lesssim 1^\circ\), this correction is
small (\(\sim 1\)km,s\(^{-1}\)), but can prevent the introduction of a
spurious gradient on the sky \citep{WMO2008}. The uncertainty is then
the velocity uncertainty plus the distance uncertainties times the PM
uncertainty from above. We then use the \(v_z\) values for the following
modelling, however repeating with plain, solar-frame velocities does not
substantially affect the results too much .

\subsection{Sculptor}\label{sculptor}

For Sculptor, we combine radial velocity measurements

\begin{itemize}
\item
  APOGEE (crossmatched to stars meeting basic quality cuts in field)
\item
  \citet{sestito+2024a}
\item
  \citet{tolstoy+2024}
\end{itemize}

We adopt a very loose prior of \[
\mu_{v} = N(111, 10) \\
\log\sigma_{v} = N(1, 1)
\] Our resulting velocity dispersions are \$\$

\$\$ Comparing to past work, note that our values are reasonable.

\subsection{Ursa Minor}\label{ursa-minor}

\begin{itemize}
\tightlist
\item
  APOGEE
\item
  \citet{sestito+2024b}
\item
  \citet{pace+2022}
\item
  \citet{spencer+2018}
\end{itemize}

\subsection{Discussion and
limitations}\label{discussion-and-limitations}

Our model here is relatively simple. Some things which we do not account
for include

\begin{itemize}
\tightlist
\item
  interstudy systematics and biases
\item
  Inappropriate uncertainty reporting.
\item
  Binarity. Binay stars are expected to induce an average velocity
  dispersion of xxx on an individual star.
\item
  Selection effects. RV studies each have their own targeted program and
\end{itemize}

We partially address several of these by also performing the analysis on
large RV samples for each galaxy. This allows for individual comparison
illustrating the differences in selection effects.

### 2025-04-15 thesis/introduction.tex
fixes compiling and warnings
thesis updates

\begin{figure}
\centering
\includegraphics{/Users/daniel/Library/Application Support/typora-user-images/image-20250408091622850.png}
\caption[Cosmological Power Spectrum]{figure 1 from
\citet{bechtol+2022}.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{figures/scl_umi_vs_penarrubia.pdf}
\caption[Idealized simulations match Scl and UMi]{Sculptor and UMi's
profiles are well-matched to \citet{PNM2008}.}\label{fig:toy_profiles}
\end{figure}

\section{}\label{section}

@@

\begin{itemize}
\tightlist
\item
  \citet{battaglia+2008}
\item
  \citet{iorio+2019}
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  \citet{sestito+2023a}
\item
  \citet{tolstoy+2023}, \citet{arroyo-polonio+2023},
  \citet{arroyo-polonio+2024}
\item
  \citet{eskridge1988}
\end{itemize}

@@

\begin{itemize}
\item
  \citet{sestito+2023b}
\item
  \citet{pace+2020}
\end{itemize}

@@

\begin{table*}[t]
\centering
\caption{Measured properties of Ursa Minor}
\label{tbl:Measured-properties-of-Ursa-Minor}
\begin{tabular}{lll}
\toprule
parameter & value & Source\\
\midrule
$\alpha$ & $ 227.2420 \pm 0.0045$˚ & M+18\\
$\delta$ & $67.2221 \pm 0.00158$˚ & M+18\\
distance & $70.1 \pm 2.9$ kpc & \\
$\mu_\alpha \cos \delta$ & $-0.124 \pm 0.004 \pm 0.017$ mas yr$^{-1}$ & MV20a\\
$\mu_\delta$ & $0.078 \pm 0.004_{\rm stat} \pm 0.017_{\rm sys}$ mas yr$^{-1}$ & MV20a\\
RV & $-245 \pm 1$ km s$^{-1}$ & This work\\
$\sigma_v$ & $9 \pm 0.6$ & This work\\
$r_h$ & $18.2 \pm 0.1$ arcmin & M+18\\
ell & $0.55 \pm 0.01$ & M+18\\
PA & $50 \pm 1^\circ$ & M+18\\
$M_V$ & $-9.03 \pm 0.05$ & M+18\\
\bottomrule
\end{tabular}
\end{table*}

@@

\begin{figure}
\centering
\includegraphics{/Users/daniel/thesis/figures/idealized_break_radius.png}
\caption[Break radius validation]{The break radius of the simulations is
set by}\label{fig:idealized_break_radius}
\end{figure}

>>>>

\begin{figure}
\centering
\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{figures/power_spectrum.png}
\caption[Cosmological Power Spectrum]{figure 1 from
\citet{bechtol+2022}.}\label{fig:cosmological_power_spectrum}
\end{figure}

@@

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_umi_vs_penarrubia.pdf}}
\caption[Idealized simulations match Scl and UMi]{Sculptor and UMi's
profiles are well-matched to \citet{PNM2008}.}\label{fig:toy_profiles}
\end{figure}

@@

\begin{itemize}
\tightlist
\item
  \citet{battaglia+2008}
\item
  \citet{iorio+2019}
\item
  \citet{amorisco+zavala+deboer2014}
\item
  \citet{battaglia+2008}
\item
  \citet{breddels+2013}
\item
  \citet{breddels+helmi2013}
\item
  \citet{richardson+fairbairn2014}
\item
  \citet{SFW2017}
\item
  \citet{innanen+papp1979}
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  \citet{sestito+2023a}
\item
  \citet{tolstoy+2023}, \citet{arroyo-polonio+2023},
  \citet{arroyo-polonio+2024}
\item
  \citet{eskridge1988}, \citet{eskridge1988a}, \citet{eskridge1988b}
\item
  \citet{coleman+dacosta+bland-hawthorn2005}
\item
  \citet{DQ1994}
\item
  \citet{WMO2009}
\item
  \citet{IH1995}, \citet{munoz+2018}
\item
  \citet{kirby+2009}
\item
  \citet{martinez-vazquez+2015}, \citet{pietrzynski+2008}
\end{itemize}

@@

\begin{itemize}
\item
  \citet{sestito+2023b}
\item
  \citet{pace+2020}
\item
  \citet{bellazzini+2002}
\item
  \citet{hargreaves+1994}
\item
  \citet{martinez-delgado+2001}
\item
  \citet{munoz+2005}
\item
  \citet{pace+2020}
\item
  \citet{palma+2003}
\item
  \citet{spencer+2018}
\item
  \citet{vitral+2023}
\end{itemize}

@@

\begin{table*}[t]
\centering
\caption{Measured properties of Ursa Minor. References are: (1) Muñoz et al. (2018), (2) Garofalo et al. (2025), (3) McConnachie and Venn (2020), (4) average of Pace et al. (2020) and Spencer et al. (2018).}
\label{tbl:Measured-properties-of-Ursa-Minor-References-are:-1-Muñoz-et-al-2018-2-Garofalo-et-al-2025-3-McConnachie-and-Venn-2020-4-average-of-Pace-et-al-2020-and-Spencer-et-al-2018}
\begin{tabular}{lll}
\toprule
parameter & value & Source\\
\midrule
$\alpha$ & $ 227.2420 \pm 0.0045$˚ & 1\\
$\delta$ & $67.2221 \pm 0.00158$˚ & 1\\
distance modulus & $19.23 \pm 0.11$ & 2\\
distance & $70.1 \pm 3.6$ kpc & \\
$\mu_\alpha \cos \delta$ & $-0.124 \pm 0.004 \pm 0.017$ mas yr$^{-1}$ & 3\\
$\mu_\delta$ & $0.078 \pm 0.004_{\rm stat} \pm 0.017_{\rm sys}$ mas yr$^{-1}$ & 3\\
RV & $-244.9 \pm 0.3_{\rm stat} \pm 1.7_{\rm sys}$ km s$^{-1}$ & 4\\
$\sigma_v$ & $8.3 \pm 0.3_{\rm stat} \pm 0.4_{\rm sys}$ & 4\\
$r_h$ & $18.2 \pm 0.1$ arcmin & 1\\
ell & $0.55 \pm 0.01$ & 1\\
PA & $50 \pm 1^\circ$ & 1\\
$M_V$ & $-9.03 \pm 0.05$ & 1\\
\bottomrule
\end{tabular}
\end{table*}

@@

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{/Users/daniel/thesis/figures/idealized_break_radius.png}}
\caption[Break radius validation]{The break radius of the simulations is
set by}\label{fig:idealized_break_radius}
\end{figure}

### 2025-04-15 thesis/main.tex
thesis updates (most of data chapter)
fixes compiling and warnings
thesis updates

\title{From tides to tails: \\
    Does the Milky Way affect Sculptor and Ursa Minor?
}
\author{Daniel A. Boyea}

@@

\addcontentsline{toc}{chapter}{Supervisory Committee}
\begin{centering}
\textbf{\@title}
\skipline
by\\
\skipline
\@author\\
B.Sc. The Ohio State University, 2023\\
\vspace*{3\baselineskip}
\end{centering}

\subsection*{Supervisory Committee}

@@

\noindent Dr. Kim Venn, Co-Supervisor\\
Department of Astronomy

@@

% \chapter*{Acknowledgements}
% \addcontentsline{toc}{chapter}{Acknowledgements}
%

@@

@@

\chapter{Model parameters}\label{sec:jack}

\chapter{Font properties}

textwidth: \the\textwidth

\newlength{\myw}
\newlength{\myl}

---\rule[2.4pt]{12pt}{0.6pt}
\settowidth{\myw}{---}\the\myw

{\bf ---}\rule[2.2pt]{12pt}{1.1pt}

\settoheight{\myl}{---}\the\myl

-\rule[2.4pt]{3pt}{0.75pt}
I\rule{1.2pt}{8pt}
{\bf I}\rule{2.0pt}{8pt}
|\rule[-2pt]{0.6pt}{12pt}

>>>>

\title{Timid tides or tall tails? \\
    Investigating the outer reaches of the Sculptor and Ursa Minor dwarf galaxies
}
\author{Daniel A. Boyea}

@@

\addcontentsline{toc}{chapter}{Supervisory Committee}
\begin{centering}
\textbf{\@title}\\
\skipline
by\\
\skipline
\@author\\
B.Sc. The Ohio State University, 2023\\
\vspace*{3\baselineskip}
\end{centering}

\subsection*{Supervisory Committee}
\skipline

@@

\noindent Dr. Kim Venn, Co-Supervisor\\
Department of Astronomy\\

\skipline

\noindent External Examiner \\
Outside Department

@@

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

%Julio Navarro, my supervisor. Rapha and other collegues

%Friends and family and collegues (fill in later).

This work has made use of data from the European Space Agency (ESA) mission
{\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia}
Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
has been provided by national institutions, in particular the institutions
participating in the {\it Gaia} Multilateral Agreement. 
\citep{gaiacollaboration+2016, gaiacollaboration+2023}.

This research made use of hips2fits,\footnote{https://alasky.cds.unistra.fr/hips-image-services/hips2fits} a service provided by CDS.

@@

\chapter{Data Selection}
\input{data}

\chapter{Methods}
\input{methods}

@@

\chapter{Model parameters}

\input{numerical_convergence}
